### 2021-04-20 input/figproof.tex
about proof, about implementation, merge complexity and implementation

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:proofA}Stale $\alpha$ messages may
      stop other $\alpha$ messages from reaching all \processes ($b$
      and $c$) that require it along the shortest path. Consistent
      partitioning requires eventual purging of stale messages (see
      Figure~\ref{fig:proofB}).]  {\input{input/figproofA.tex}}
    \hspace{10pt}
    \subfloat[Part B][\label{fig:proofB}Stale $\alpha$ messages may
      stop $\delta$ messages from reaching processes with the corresponding
      partition. Consistent partitioning requires that all \processes propagate
      $\delta$ messages as often as possible (I'); detect possible inconsistencies
      when parents' $\alpha$ messages contradict delivery history (III); notify
      processes of possible inconsistencies (IV).]
    {\input{input/figproofB.tex}}
    \caption{\label{fig:proof}Concurrent operations,
      delivery orders, and stale control information may impair
      consistent partitioning. Dynamic consistent partitioning
      requires eventual purging of stale messages, and best eventual
      forwarding.}
  \end{center}
\end{figure*}

>>>>

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:proofA}Stale $\alpha$ messages may
      stop other $\alpha$ messages from reaching all \processes ($b$
      and $c$) that require it along the shortest path. Consistent
      partitioning requires eventual purging of stale messages (see
      Figure~\ref{fig:proofB}).]  {\input{input/figproofA.tex}}
    \hspace{10pt}
    \subfloat[Part B][\label{fig:proofB}Stale $\alpha$ messages may
      stop $\delta$ messages from reaching processes with the corresponding
      partition. Consistent partitioning requires that all \processes propagate
      $\delta$ messages as often as possible (I'); detect possible inconsistencies
      when parents' $\alpha$ messages contradict delivery history or state (III); notify
      processes of possible inconsistencies (IV).]
    {\input{input/figproofB.tex}}
    \caption{\label{fig:proof}Concurrent operations,
      delivery orders, and stale control information may impair
      consistent partitioning. Dynamic consistent partitioning
      requires eventual purging of stale messages, and best eventual
      forwarding.}
  \end{center}
\end{figure*}

### 2021-04-20 input/figproofB.tex
purging
about proof, about implementation, merge complexity and implementation

  \draw[->] (0 +5pt, 0)
  node[below=-0.3em, below right, font=\tiny]{$\textcolor{\PF}{\ADD_f^{z'}} \rightarrow$}
  --
  (\X -5pt, 0)
  node [left= -0.15em, below=-0.3em, below left, font=\tiny]{$\textcolor{\WRONG}{\DEL_f \not\rightarrow \ADD_f^{z'}}$}
  ; %% e - b'

@@

  \draw[->] ( \X +5pt, 0) -- (2*\X -5pt, 0); %% b - b''

  \draw[fill=white] (-\X, 0) node{$\bm{d}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);  
  \draw[fill=white] (0, 0) node[color=\PF]{$\bm{e}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (\X, 0) node[color=\PD]{$\bm{b'}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (2*\X, 0) node[color=\PD]{$\bm{b''}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);

@@

  \draw ( 0, 5pt) node[above, font=\footnotesize]{$
    \xrightarrow[\textcolor{\PC}{\ADD_d^{\vphantom{z'}}}]{\mbox{\scriptsize{before}}}
    \xrightarrow[\textcolor{\PF}{\ADD_f^{z\vphantom{'}}}]{\mbox{\scriptsize{last}}}
    \xrightarrow[\DEL_f^{\vphantom{z'}} ]{\mbox{\scriptsize{expect}}}$};

  \draw ( \X, 5pt) node[above, font=\footnotesize]{$
    \xrightarrow[\textcolor{\PF}{\ADD_f^{\vphantom{y'}}} \rightarrow \DEL_f]{\mbox{\scriptsize{before}}}
    \xrightarrow[\textcolor{\PC}{\ADD_d^{y'}}]{\mbox{\scriptsize{last}}}
    \xrightarrow[\DEL_d^{\vphantom{z'}}?]{\mbox{\scriptsize{expect}}}$};

@@

  \draw[->] ( 0,  0.5*\Y) -- ( 0, -5pt);
  \draw[dotted] ( 0,  0.5*\Y) -- ( 0, 0.9*\Y) node[below right, font=\scriptsize]{somewhere: $\textcolor{\PF}{\ADD_f} \rightarrow \DEL_f$};

>>>>

  \draw[->] (0 +5pt, 0)
  node[below=-0.3em, below right, font=\tiny]{$\textcolor{\PH}{\ADD_h^{z'}} \rightarrow$}
  --
  (\X -5pt, 0)
  node [left= -0.15em, below=-0.3em, below left, font=\tiny]{$\textcolor{\WRONG}{\DEL_h \not\rightarrow \ADD_h^{z'}}$}
  ; %% e - f

@@

  \draw[->] ( \X +5pt, 0) -- (2*\X -5pt, 0); %% b - g

  \draw[fill=white] (-\X, 0) node{$\bm{d}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);  
  \draw[fill=white] (0, 0) node[color=\PH]{$\bm{e}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (\X, 0) node[color=\PD]{$\bm{f}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);
  \draw[fill=white] (2*\X, 0) node[color=\PD]{$\bm{g}$} +(-5pt, -5pt) rectangle +(5pt, 5pt);

@@

  \draw ( 0, 5pt) node[above, font=\footnotesize]{$
    \xrightarrow[\textcolor{\PC}{\ADD_d^{\vphantom{z'}}}]{\mbox{\scriptsize{before}}}
    \xrightarrow[\textcolor{\PH}{\ADD_h^{z\vphantom{'}}}]{\mbox{\scriptsize{last}}}
    \xrightarrow[\DEL_h^{\vphantom{z'}} ]{\mbox{\scriptsize{expect}}}$};

  \draw ( \X, 5pt) node[above, font=\footnotesize]{$
    \xrightarrow[\textcolor{\PH}{\ADD_h^{\vphantom{y'}}} \rightarrow \DEL_h]{\mbox{\scriptsize{before}}}
    \xrightarrow[\textcolor{\PC}{\ADD_d^{y'}}]{\mbox{\scriptsize{last}}}
    \xrightarrow[\DEL_d^{\vphantom{z'}}?]{\mbox{\scriptsize{expect}}}$};

@@

  \draw[->] ( 0,  0.5*\Y) -- ( 0, -5pt);
  \draw[dotted] ( 0,  0.5*\Y) -- ( 0, 0.9*\Y) node[below right, font=\scriptsize]{somewhere: $\textcolor{\PH}{\ADD_h} \rightarrow \DEL_h$};

### 2021-04-20 input/proposal.tex
about proof, about implementation, merge complexity and implementation

\begin{definition}[\label{def:lwo}Last win order]
  When a \process $p$ broadcasts two messages $b_p(m) \rightarrow
  b_p(m')$, no \process $q$ can deliver $m$ if it has delivered $m'$:
  $\not\exists q \in V$ with $d_q(m') \rightarrow d_q(m)$, for $m'$
  convey \emph{stale} control information.
\end{definition}

@@

\begin{theorem}[Propagations $+$ detection $\implies$ purge]
  \Processes can purge stale control information by broadcasting a
  dedicated notification message $\delta$, and detect possible
  inconsistencies using their history to propagate 
\end{theorem}

\begin{proof}
  Figure~\ref{fig:proofB} depicts all possible cases that need
  appropriate purging. A chain of \processes $[d, e, b', b'']$
  delivered and forwarded a message $\alpha_d$. Some \processes may
  have delivered messages about other partitions before or after
  $\alpha_d$.
  \begin{asparadesc}
  \item [\processes with last $\alpha_d$:] \Processes such as
    \Process~$d$ where the last and best partition is still $P_d$. By
    propagating from process to process the corresponding $\delta_d$
    message, these \processes purge $\alpha_d$.
  \item [\processes with $\alpha_d^y \rightarrow \alpha_f^z$ with last
    $\alpha_f^z$, for $\alpha_f^z < \alpha_d^y$:] \Processes such as
    \Process~$e$ do not directly suffer from the non-delivery of
    $\delta_d$. They need to purge their own $P_f$ if need
    be. Nonetheless, they deliver and forward $\alpha_f$ that at least
    one following \process will receive:
  \item [\processes with $\delta_f \rightarrow \alpha_d$:] \Processes
    such as \Process~$b'$ do not belong to the preceding category, for
    they already delivered $\delta_f$. However, best eventual
    forwarding guarantees that they receive $\alpha_f$ that
    contradicts their history or state. In such case, the detecting
    \processes purge their own $\alpha_d$. The detecting \processes
    must also notify subsequent \processes that their current
    partition \emph{may be} deleted, for their parent may have blocked
    the corresponding $\delta$ messages. We note such notification
    $\delta_d?$ to emphasize the uncertainty of a $\delta$ message.
  \item [\processes with last $\alpha_d$ receiving $\delta_d?$ from
    their parent:] \Processes such as \Process~$b''$ have $\alpha_d$
    and preceding delivered messages have not importance. Such
    \processes must trust detecting \processes by delivering and
    forwarding $\delta_d?$. Messages are subject to all aforementioned
    blocking conditions, but are also solved by aforementioned
    mechanisms.
  \end{asparadesc}

  Propagation I' terminates: a process do not deliver a $\delta$
  message if it has already delivered it. Propagation IV terminates: a
  process do not deliver a looping message.  Propagation and detection
  ensure that \processes eventually purge stale $\alpha$ messages from
  the system, leaving room for Propagation I.
  %% In the meantime, \processes that still belong to a partition can
  %% propagate their respective last $\alpha$ message, and reach
  %% consistent partitioning.
\end{proof}

\subsection{Implementation}

Algorithm~\ref{algo:ascast} provides the instructions that
implement dynamic consistent partitioning:
\begin{inparaenum}[(i)]
\item \label{algo:most} most of the time, it propagates $\alpha$'s and
  $\delta$'s when the partitions allow it;
\item \label{algo:sometimes} sometimes, it detects possible
  inconsistent partitioning and
\item \label{algo:solves}solves it using propagation trees (as opposed
  to propagation graphs);
\item \label{algo:competition} when required, it triggers competitions
  among neighboring partitions.
\end{inparaenum}
While (\ref{algo:most})-(\ref{algo:solves}) tackle the eventual purging of
stale notifications, (\ref{algo:competition}) tackle the eventual
delivery of the best up-to-date partitions.

@@

To constrain the order of delivery at each \process, each \process
implements last win order by maintaining a vector of versions that
associates to each known source, or has-been source, with its known
local counter. It enables \processes to deliver only up-to-date
$\alpha$'s that may improve their best known partition (see
Line~\ref{line:detectA}), but also to detect possible inconsistent
partitioning (see Line~\ref{line:detectB}) as labeled by III in
Figure~\ref{fig:proofB}. From a global perspective, it ensures
monotonic increase of knowledge despite the non-monotonic number of
sources. In turns, it contributes to convergence and termination, for
corresponding $\alpha$ and $\delta$ cannot follow each other in an
infinite loop.

Implementing the propagation of deletes is no different from the
propagation of adds as hinted in Figure~\ref{fig:proof} with Labels I
and I'. Messages and comparisons include the local counter of the
source. When a \process receives a better up-to-date partition, it
delivers and forwards it. When a \process receives a delete
notification about its partition, it delivers and forwards it. It is
worth noting that such $\delta$'s apply to the whole partition. They
travel in the propagation graph of their partition using all
communication links, allowing \processes to change partition as soon as
possible.

Vectors of versions also allow each \process to detect possible last
win order violations. Upon possible inconsistency, the detecting
\process must remove all $\alpha$ messages originating from it, as
labeled by IV in Figure~\ref{fig:proofB}. Contrarily to delete
operations, this applies to a sub-partition of the current partition,
and cannot rely on local counters to
operate. Algorithm~\ref{algo:ascast} implements this behavior by
piggybacking paths in messages. Paths guarantee that such $\delta$
messages apply to all and only $\alpha$'s that followed an identical
path. $\delta$'s propagate in the propagation tree rooted at the
detecting \process. Local states of \processes remain unchanged, but
paths in messages monotonically grow. In turns, $\alpha$ and $\delta$
cannot follow each other in an infinite loop, for a \process knows when
it already received a forwarded a message (see
Lines~\ref{line:notloopingA} and \ref{line:notloopingB}).

\noindent It is worth noting that scoped broadcast and the
piggybacking of paths synergies well: Paths tend to be smaller as the
number of sources in the system increases.

Finally, as stated in Theorem~\ref{theo:dcp}, dynamic consistent
partitioning not only requires the removal of all stale partitions,
but also the retrieval of the best up-to-date partitions. In that
regard, $\delta$'s have dual use: since they already reach borders of
partitions when they remove stale control information, they also
trigger a competition when reaching such bordering \processes (see
Line~\ref{line:compete}). This simply consists in sending the current
partition through the communication link from which the \process
received the $\delta$. Upon receipt of this answer, \processes act
normally by propagating their changes when they improve. These adds
fill the gaps left open by deletes.

@@

Algorithm~\ref{algo:edges} provides the instructions of \NAME that
enable dynamic consistent partitioning in dynamic networks where
\processes can join or leave the system without
notification. \TODO{More.}

%% Adding new communication links to the network may create shortcuts
%% between \processes. Both \processes must send their current best
%% partition to each other. Upon receipt, they act normally: if a \process
%% finds out that the received partition is closer than its current one,
%% it delivers it which in turns also triggers another competition
%% amongst neighbors due to forwarding.

%% \noindent Joining the network is equivalent to add as many
%% communication links as necessary between the joining \process and its
%% new neighbors.

%% When removing a communication link between two \processes does not
%% break any active path, because neither distances of \processes depend
%% on the other, then nothing needs to be done. \NAME has no overhead.
%% Unfortunately, when a \process' distance depends on the other \process,
%% the protocol becomes much more complex. Indeed, this requires to
%% \TODO{undo} all add messages originated from this \process. A message
%% must convey the fact that
%% \begin{inparaenum}[(i)]
%% \item an edge at a particular \process has been removed, and
%% \item the distance that has been delivered by a \process comes from
%%   this particular \process.
%% \end{inparaenum}

%% \subsection{Inter-autonomous system partitioning}

%% \begin{figure}
%%   \centering\input{input/figAS.tex}
%%   \caption{\label{fig:AS}Inter autonomous systems
%%     partitioning. \TODO{More about relative ordering of links at each
%%       \process}}
%% \end{figure}

%% In this section, we aim at solving Problem Statement~\ref{prob:inter}
%% by proposing \NAMEB, an extension of \NAME, that enables dynamic
%% logical partitioning in inter-autonomous systems. It leverages
%% hierarchical properties of these networks to further improve on scoped
%% flooding.

%% Autonomous systems are geo-distributed networks. Heterogeneous
%% communication links. Hierarchy of communication links. Per-object
%% broadcast.

%% We leave the link handling to membership protocols~\REF. The ordering
%% of links can be done based on latency. For instance, from 0 to 100 ms,
%% rank 0, from 100 to 1s, rank 2 \ldots

%% \begin{algorithm}
%%   \input{input/algoaaaa.tex}
%%   \caption{\label{algo:aaaa}\NAMEB running at \Process $p$ for dynamic
%%     partitioning in inter-autonomous systems.}
%% \end{algorithm}

%% Figure~\ref{fig:AS} shows an example of inter-AS logical partitioning
%% where each \process ranks its communication links, and each \process
%% forwards messages to its neighbors starting from the rank of the link
%% that triggered the receipt, to its highest rank neighbors. In this
%% example, \Process $e$ adds a partition. It forwards the generated
%% message to all its neighbors as if it received it from its lowest rank
%% links. When \Process $a$ receives $\alpha_e^1$ from its rank-$0$ link,
%% it forwards it to rank-$0$ links and rank-$1$ links.  \Process $b$
%% stops the propagation, for all its links have lower ranks than the one
%% from which it received the message $\alpha_e^2$. Ultimately, the
%% partition includes \Process $e$, \Process $a$, and \Process $b$. \Process
%% $c$ and \Process $d$ remain unaware of the new \TODO{object}.

%% To ease the reasoning about inter-AS dynamic partitioning in dynamic
%% networks, \NAME runs an independent broadcast protocol for each
%% object, and for each rank of links.  \TODO{Not so easy\ldots}
%% \TODO{Write and describe algo.} \TODO{Overhead of separating things,
%%   slightly more memory, but traffic wise? slower ?}  \TODO{In the
%%   example, not only \Process $b$ keeps $e_1$ as its best partition, but
%%   it keeps it for links of rank 0, and links of rank 1.}  When there
%% are no ranking in links, \processes solely rely on
%% Algorithm~\ref{algo:adddelundo} for optimal partitioning.

\subsection{Complexity}

We focus on average-case and worst-case complexity. We divide our
analysis into space, time, and communication complexity.

\begin{asparadesc}
\item [Space:]

Each \process needs to maintain a vector of versions the size of which
depends on the number of sources. In the worst case, the size of this
vector increases linearly with the number of distinct sources that
ever existed in the system. For the sake of last win order, \processes
can never purge this vector from any entry. \TODO{Even when the
  \process leaves the system.} Fortunately, scoped broadcast limits
message propagation to interested \processes. A \process at a side of
the system may never hear of a source at the other side of the
system. In turns, it may never have to maintain an entry for this
source. \TODO{This is application-dependent?}

\noindent Each \process must also maintain the partition it belongs to,
along with its control information. Among other, it must maintain a
path the size of which depends on the membership size of the
partition. In the worst case, this path includes the identity of every
\process of the system that currently belong to the system. Eventually,
the size of this path increases linearly with the diameter of the
current partition. \TODO{assuming weight equals 1 \ldots}

\item[Time:]

In terms of processing time, whatever the operation, it increases
linearly with the number of neighbors of the \process. It scales well,
for the number of neighbors is commonly order of magnitude lower than
the number of \processes in the system. For instance, random networks
only require $O(\log(|V|))$ neighbors to work properly \REF; and
\processes in autonomous systems have in average \TODO{X} neighbors
\REF. Lines~\ref{line:notloopingA} and \ref{line:notloopingB} can
prove costly depending on the data structure encoding paths. Using
Bloom filters however, this is \TODO{efficient}: it depends on the
number of buckets, and the number of hash functions.

\noindent In terms of convergence time, every $\alpha$ needs to travel
at least its shortest path before termination. Every $\delta$ needs to
reach the border of its partition, triggering an $\alpha$ that fills
the gaps left open by the removal. \TODO{More.}

\item[Communication:]

In terms of number of messages required to reach optimal
partitioning. In the average-case, a \process $i$ chosen uniformly at
random among all \processes creates a logical partition. Its messages
$\alpha_i$ propagate through the network until reaching \processes
that belong to another partition closer to them. This splits
partitions in half in average. Overall, the $a^{th}$ new partition
comprises \smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor \log_2 a
      \rfloor}})$} \processes. This decreases every new partition
until reaching $0$ \processes per new partition: even the chosen
\process already belongs to its optimal partition. The average number
of messages per \process is
\smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor \log_2 a
      \rfloor}})$}. \TODO{Multiple receipt and multiple delivery imply
  more messages (receipt bounded by $|O|$ as well).} Deleting the
$a^{th}$ partition generates the exact same number of messages than
the $a^{th}$ partition creation. \TODO{But what about echos?} In the
worst-case, every new partition includes all but one \process
belonging to the previous partition. The total number of messages
after the $a^{th}$ new partition is $\mathcal{O}(\overline{|O|}\cdot
a^2)$. As for the average-case, the number of messages for the
partition deletion is identical to the number of messages of the
corresponding partition creation.

\noindent \TODO{In terms of size of messages.}

\end{asparadesc}

>>>>

\begin{definition}[\label{def:lwo}Last win order and staleness]
  When a \process $p$ broadcasts two messages $b_p(m) \rightarrow
  b_p(m')$, no \process $q$ can deliver $m$ if it has delivered $m'$:
  $\not\exists q \in V$ with $d_q(m') \rightarrow d_q(m)$, for $m$
  convey \emph{stale} control information.
\end{definition}

@@

\begin{theorem}[three-phase purge] % Propagations $+$ detection $\implies$ purge]
  Purging stale control information from its source to every \process
  that may have delivered it last requires three
  phases:
  \begin{inparaenum}[(i)]
  \item propagation of delete notifications,
  \item detection of possible blocking conditions, and
  \item propagation of possibly deleted but blocked notifications.
  \end{inparaenum}
\end{theorem}

\begin{proof}
  Figure~\ref{fig:proofB} depicts all possible cases that need
  appropriate purging. A chain of \processes $[d, e, f, g]$ delivered
  and forwarded a message $\alpha_d$. Some \processes may have
  delivered messages about other partitions before or after
  $\alpha_d$.
  \begin{asparadesc}
  \item [\processes with last $\alpha_d$:] \Processes such as
    \Process~$d$ where the last and best partition is still $P_d$. By
    propagating from \process to \process the corresponding $\delta_d$
    message, these \processes purge $\alpha_d$.
  \item [\processes with $\alpha_d^y \rightarrow \alpha_h^z$ with last
    $\alpha_h^z$, for $\alpha_h^z < \alpha_d^y$:] \Processes such as
    \Process~$e$ do not directly suffer from the non-delivery of
    $\delta_d$. They need to purge their own $P_h$ if need
    be. Nonetheless, they deliver and forward $\alpha_h$ that at least
    one following \process will receive:
  \item [\processes with $\delta_h \rightarrow \alpha_d$:] \Processes
    such as \Process~$f$ do not belong to the preceding category, for
    they already delivered $\delta_h$. However, best eventual
    forwarding guarantees that they receive $\alpha_h$ that
    contradicts their history or state. In such case, the detecting
    \processes purge their own $\alpha_d$. The detecting \processes
    must also notify subsequent \processes that their current
    partition \emph{may be} deleted, for their parent may have blocked
    the corresponding $\delta$ messages. We note such notification
    $\delta_d?$ to emphasize the uncertainty of a $\delta$ message.
  \item [\processes with last $\alpha_d$ receiving $\delta_d?$ from
    their parent:] \Processes such as \Process~$g$ have $\alpha_d$ and
    preceding delivered messages have not importance. Such \processes
    must trust detecting \processes by delivering and forwarding
    $\delta_d?$. Messages are subject to all aforementioned blocking
    conditions, but are also solved by aforementioned mechanisms.
  \end{asparadesc}

  Propagation~I' terminates: a \process do not deliver a $\delta$
  message if it has already delivered it. Propagation~IV terminates: a
  \process do not deliver a looping message.
  % ; and since \processes forward the messages they deliver, upstream
  % \processes eventually receive and may deliver messages detected as
  % issue. In Figure~\ref{fig:proofB}, \Process~$e$ eventually
  % receives $\delta_f$ from \Process~$b'$.
  These three phases propagation-detection-propagation ensure that
  \processes eventually purge stale $\alpha$ messages from the system,
  leaving room for Propagation~I.
  %
  %  \TODO{cannot loop add from $e$ , undo from $b'$ because $b'$ sent
  % delete to $e$, it will solve the inconsistency.}
  %
  %% In the meantime, \processes that still belong to a partition can
  %% propagate their respective last $\alpha$ message, and reach
  %% consistent partitioning.
\end{proof}

\subsection{Implementation and complexity}

Algorithm~\ref{algo:ascast} provides the few instructions of \NAME
that implement three-phase purge and best eventual forwarding to
enable dynamic consistent partitioning.

%% Similarly to Algorithm~\ref{algo:add}, local operations \texttt{Add}
%% and \texttt{Del} generate messages that the initiating \process treats
%% identically to receipts of messages from remote \processes.

%% \begin{inparaenum}[(i)]
%% \item \label{algo:most} most of the time, it propagates $\alpha$'s and
%%   $\delta$'s when the partitions allow it;
%% \item \label{algo:sometimes} sometimes, it detects possible
%%   inconsistent partitioning and
%% \item \label{algo:solves}solves it using propagation trees (as opposed
%%   to propagation graphs);
%% \item \label{algo:competition} when required, it triggers competitions
%%   among neighboring partitions.
%% \end{inparaenum}
%% While (\ref{algo:most})-(\ref{algo:solves}) tackle the eventual purging of
%% stale notifications, (\ref{algo:competition}) tackle the eventual
%% delivery of the best up-to-date partitions.

@@

To implement last win order as stated in Definition~\ref{def:lwo},
each \process maintains a vector of versions that associates the
respective known local counter of each known source, or has-been
source.  Each \process delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
This vector enables both Propagation~I and Propagation~I'.  The size
of such vector increases linearly with the number of sources that the
\process ever known which is the number of \processes in the system
$\mathcal{O}(V)$ in the worst case. Unfortunately, \processes cannot
remove any entry from this vector without running a costly distributed
consensus protocol to ensure that no \process can mistake an old but
removed entry for a new one. However, using \NAMEB, we expect that
every \process knows only a small subset of sources.

This vector of versions constitutes a summary of deliveries. \NAME
uses it to detect possible inconsistent partitioning as depicted in
Figure~\ref{fig:proofB}. A detecting \process then propagates a
$\delta$ message conveying the unique identifier $s$ and version $c$
of the partition along with a path $\pi$ comprising the identity of
\processes that forwarded the corresponding $\alpha$ message. Such
$\delta$ message implements the uncertain $\delta?$ message of
Figure~\ref{fig:proofB}. It fulfils the double role of removing stale
control information without modifying the local counter of the source
(since the partition may still exist), and removing loops to ensure
termination (since \processes do not modify their local state to
reflect such delivery, messages piggyback monotonically such
increasing control information). Thus, $\delta$ messages used in
Propagation~I' are more efficient both in propagation time and
traffic, for they propagate using every communication link available
in the partition, and they do not need to carry their path.
Fortunately, Propagation~I' happens more often than Propagation~IV.
Yet, \processes that belong to a partition need to save the path taken
by the corresponding $\alpha$ messages to enable Propagation~IV.  In
the worst case, the path includes every \processes in the system but
converges in average to the average diameter of partitions. Thus,
message paths and \NAME have an interesting synergy: Paths tend to be
smaller as the number of sources in the system increases.

As stated in Theorem~\ref{theo:dcp}, dynamic consistent partitioning
not only requires eventual purging of stale control information, but
also the retrieval of the best up-to-date partitions. In that regard,
$\delta$ messages have dual use: since they already reach the borders
of their partition when they remove stale control information, they
also trigger a competition when reaching such bordering \processes
(see Line~\ref{line:ascast_compete}). This simply consists in sending
the current partition through the communication link from which the
\process received the $\delta$. Upon receipt of this answer,
\processes act normally by propagating their changes when they
improve.

\TODO{About communication complexity} In terms of number of messages
required to reach optimal partitioning. In the average-case, a
\process $i$ chosen uniformly at random among all \processes creates a
logical partition. Its messages $\alpha_i$ propagate through the
network until reaching \processes that belong to another partition
closer to them. This splits partitions in half in average. Overall,
the $a^{th}$ new partition comprises
\smash{$\mathcal{O}(\frac{|V|}{2^{\lfloor \log_2 a \rfloor}})$}
\processes. This decreases every new partition until reaching $0$
\processes per new partition: even the chosen \process already belongs
to its optimal partition. The average number of messages per \process
is \smash{$\mathcal{O}(\frac{\overline{|O|}}{2^{\lfloor \log_2 a
      \rfloor}})$}. \TODO{Multiple receipt and multiple delivery imply
  more messages (receipt bounded by $|O|$ as well).} Deleting the
$a^{th}$ partition generates the exact same number of messages than
the $a^{th}$ partition creation. \TODO{But what about echos?} In the
worst-case, every new partition includes all but one \process
belonging to the previous partition. The total number of messages
after the $a^{th}$ new partition is $\mathcal{O}(\overline{|O|}\cdot
a^2)$. As for the average-case, the number of messages for the
partition deletion is identical to the number of messages of the
corresponding partition creation.
\noindent \TODO{In terms of size of messages.}

@@

Algorithm~\ref{algo:edges} extends the instructions of \NAME that
enable dynamic consistent partitioning in dynamic networks where
\processes can join or leave the system without notification. We
consider that removing a \process is equivalent to removing all its
edges, and adding a \process is equivalent to add as many edges as
necessary.

Adding an edge between two \processes may only improve the shortest
path of one of these \processes. Therefore, it triggers a competition
between the two \processes only. Removing an edge between two
\processes may invalidate the shortest path of one of involved
\processes plus downstream \processes. To implement this behavior,
\NAME uses the implementation of $\delta?$ messages. This prove costly
but enables \NAME even in dynamic networks subject to physical
partitioning.
